---
title: "Montana NLCD Zonal Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Work flow of this Rmarkdown document:
1. Read in packages and data
2. Load data:
- NLCD raster
- Montana outline shapefile
- Montana Counties shapefile
- Stranded Land shapefile

3. Crop Stranded Land Shapefile to Montana Outline
4. Crop NLCD raster to montana outline extent and resample to coarser resolution (currently 90m)
5. Mask resampled NLCD to Montana outline
6. Reclassify NLCD -- see code chunk for specifics
7. Clean stranded land geometries (much faster in QGIS)
8. Crop NLCD to stranded land

If interested in looking at a particular county:
9. Crop MT NLCD to county outline
10. Crop stranded NLCD to that county
11. Plot to check and visualize

12. Write output shapefiles

### 1. Load packages and data
```{r packages_and_data}

library(tidyverse)
library(raster)
library(rgdal)
library(sf)
library(cleangeo)
library(parallel)

# setwd("~/StrandedLand/stranded_repo") # Bren R server
# setwd("~/Desktop/StrandedLand/stranded_repo") # Desktop

# output_resolution = 90
output_resolution = 270 # Make sure this matches with RMD 3
```

### 2. Load Data
```{r load_data}

# NLCD: Lower 48 NLCD raster data (2016)
nlcd <- raster("../R_input_files/nlcd/NLCD_2016_Land_Cover_L48_20190424.img")

# Shapefile: State of Montana outline
mt_outline <- readOGR("../R_input_files/mt_outline/StateofMontana.shp") %>% 
  # Put into NLCD CRS so the layers overlap 
  spTransform(.,crs(nlcd)) 

# Shapefile to simple features. Used later with stranded land layer
mt_sf = st_as_sf(mt_outline)
  
# Shapefile of Montana County outlines
mt_counties <- readOGR("../R_input_files/mt_counties/County.shp") %>% 
  # Put into NLCD CRS so the layers overlap 
  spTransform(.,crs(nlcd)) 

# Shapefile: Strande Land
stranded_sf <- read_sf("../R_input_files/stranded_shp/mt_strandedMaster_clean.shp") %>% 
  st_transform(.,crs(nlcd)) %>% 
  st_simplify() %>% 
  filter(stranded == 1)

```

### 3. Crop Stranded Land shapfile to the outline of Montana (spans most of the Western US)
```{r crop_stranded_land}

# Crop the stranded shapefile to Montana 
mt_stranded = st_intersection(stranded_sf,mt_sf) %>% as(.,"Spatial")

```

### 4. Crop NLCD to extent of Montana outline, then resample to the output_resolution (defined in first code chunk)
```{r crop_resample_nlcd}
# Crop NLCD to Montana Outline (shrink extent)
nlcd_crop = crop(nlcd,extent(mt_outline))

## Create raster object to resample NLCD with
# Set extent equal to cropped NLCD created above
rsmpl_ras <- raster(nlcd_crop)
rsmpl_ext <- extent(nlcd_crop)
extent(rsmpl_ras) <- rsmpl_ext

# Set resolution
res(rsmpl_ras) <- output_resolution

# Resample NLCD mt to desired resolution
nlcd_resample = resample(nlcd_crop, rsmpl_ras, method = "ngb")

# Write output raster
writeRaster(nlcd_resample,paste0("../R_output_files/nlcd/mt_nlcd_rsmpl_",output_resolution,".tif"),overwrite=TRUE)
```

### 5. Mask resampled NLCD to MT outline
```{r nlcd_mt_mask}

# Define core cluster size
n_cores <- detectCores()-6 

timestamp() # For reference

# Initiate use of core cluster
beginCluster(n_cores) 
# Resample
nlcd_mt <- clusterR(nlcd_resample, mask, args=list(mt_outline))
# Stop used of core cluster
endCluster()

timestamp() # For reference

# Write output raster
writeRaster(nlcd_mt,paste0("../R_output_files/nlcd/mt_nlcd_",output_resolution,".tif"),overwrite=TRUE)
```

### 6. Reclassify NLCD Raster to simplify classes
```{r nlcd_reclass}

rcl_mat = c(20,22.5,20, # <50% developed: Low intensity
            22.5,24,25, # 50 - 100% developed: High intensity
            30,32,30, # Barren
            40,44,40, # Forest
            50,53,50, # Shrubland
            70,74,70, # Grassland
            80,83,80, # Cultivated
            89,96,90) # Wetlands

# Reclassify Montana NLCD raster
timestamp()

beginCluster(n_cores)
nlcd_mt_rcl <- clusterR(nlcd_mt, reclassify,args = list(rcl_mat,include.lowest=TRUE))
endCluster()

timestamp()
```

### 7. Crop MT NLCD to just Stranded Land
```{r crop_mtNLCD_toStranded}
timestamp() #### 

beginCluster(n_cores) #
# Mask Montana NLCD to just Stranded Land in Montana
nlcd_mt_stranded <- clusterR(nlcd_mt_rcl, mask, args=list(mt_stranded))
endCluster()

timestamp() #### 
```

### 9. Crop MT NLCD to outline of county of interest -- needs updating
```{r crop_mtNLCD_toGB}

# beginCluster(n_cores)
# # Crop Montana NLCD to GB extent
# nlcd_gb_crop = crop(nlcd_mt_rcl,extent(gb_outline))
# # Mask nlcd_gb_crop to GB outline
# nlcd_gb <- clusterR(nlcd_gb_crop, mask, args = list(gb_outline))
# endCluster()

```

### 10. Crop GB NLCD to just stranded land in GB
```{r crop_gbNLCD_toStranded}
# 
# beginCluster(n_cores)
# # Crop GB NLCD to just stranded land in GB
# nlcd_gb_stranded <- clusterR(nlcd_gb, mask, args = list(gb_stranded))
# endCluster()

```

### 11. Plot all to check overlap
```{r check_overlap}

# # Montana Overview
# plot(nlcd_mt)
# plot(nlcd_mt_stranded,col="red",add=TRUE)
# plot(nlcd_gb,col=topo.colors(10),add=TRUE)
# 
# # Gallatin-Broadwater
# plot(nlcd_gb)
# plot(nlcd_gb_stranded, col = "red", add=TRUE)

```

### 12. Write Output Files
```{r write_output}

# dir.create("./output/stranded_shp/montana", recursive = T)
# dir.create("./output/stranded_shp/gb")
# dir.create("./output/nlcd/")

writeOGR(obj = mt_stranded,
         dsn = "../R_output_files/stranded_shp/montana/",
         layer = "mt_stranded.shp",
         driver = "ESRI Shapefile",
         overwrite_layer = TRUE)

# Montana NLCD
writeRaster(nlcd_mt_rcl,paste0("../R_output_files/nlcd/mt_nlcd_rcl_",output_resolution,".tif"),overwrite=TRUE)

# Montana stranded NLCD
writeRaster(nlcd_mt_stranded,paste0("../R_output_files/nlcd/mt_nlcd_rcl_strnd_",output_resolution,".tif"),overwrite=TRUE)
```

