---
title: "3_CS_Resistance_Layers"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages and Working Directory
```{r load_packages}
library(raster)
library(tidyverse)
library(sf)
library(rgdal)
library(rgeos)
library(spatialEco)
library(parallel)

# setwd("~/StrandedLand/stranded_repo") # Bren R server
# setwd("~/Desktop/StrandedLand/stranded_repo") # Desktop

# output_resolution = 90
output_resolution = 270 # Make sure this matches with RMD 1
```

## Montana/County Outlines and NLCD data
```{r mt_outlines_and_NLCD}

# Read in orignial NLCD layer to use CRS for all other layers
nlcd <- raster("../R_input_files/nlcd/NLCD_2016_Land_Cover_L48_20190424.img")

# Get MT NLCD raster data, developed in Rmarkdown document 1_layer_processing.Rmd
mt_nlcd = raster(paste0("../R_output_files/nlcd/mt_nlcd_",output_resolution,".tif"))
# Change cells with 0 value to NA
values(mt_nlcd)[values(mt_nlcd)==0] <- NA

# Get MT counties shapefile
mt_counties <- readOGR("../R_input_files/mt_counties/County.shp")%>% 
  spTransform(.,crs(nlcd))

# Get MT outline shapefile
mt_outline <- readOGR("../R_input_files/mt_outline/StateofMontana.shp") %>%
  spTransform(.,crs(nlcd))
```

## DEM/Slope

* Read in files
* Merge into one raster layer
* put into mt_nlcd CRS
* Crop to mt_outline
* Mask to mt_outline

```{r get_dem_convert2slope}

# timestamp()
# dem_list <- list.files(path = "../R_input_files/dem",recursive = T,full.names = T)
# dem_rasters <- lapply(dem_list, raster)
# 
# timestamp()
# # Merge each dem file into one raster layer
# dem <- do.call(merge,dem_rasters) # This takes ~12-15 minutes
# 
# timestamp()
# # Reproject dem
# dem_proj = projectRaster(dem,crs = crs(nlcd)) # This takes ~3 hrs?
# writeRaster(dem_proj,"../R_output_files/DEM/dem_reproj.tif", overwrite = TRUE)
# 
# dem_proj <- raster("../R_output_files/DEM/dem_reproj.tif")
# 
# timestamp() # 14:39
# # Crop to mt outlines
# dem_crop = crop(dem_proj,mt_outline)
# # writeRaster(dem_crop,"../R_output_files/DEM/dem_crop.tif", overwrite = T)
# 
# timestamp() # 14:40
# 
# beginCluster(10)
# dem_mask = clusterR(dem_crop,mask,args = list(mt_outline))
# endCluster()
# writeRaster(dem_mask,"../R_output_files/DEM/dem_mt_mask.tif", overwrite = T)
# 
# timestamp() # 14:46
# 
# 
# mt_slope = terrain(dem_mask, opt = 'slope', unit='degrees')
# writeRaster(mt_slope,"../R_output_files/DEM/slope.tif", overwrite = T)
# 
# 
# mt_slope <- raster("../R_output_files/DEM/slope.tif")
# 
# timestamp()
# mt_slope_rsmpl <- resample(mt_slope,mt_nlcd)
# writeRaster(mt_slope_rsmpl,paste0("../R_output_files/DEM/mt_slope_",output_resolution,".tif"), overwrite = T)
# writeRaster(mt_slope_rsmpl,paste0("../R_output_files/circuitscape/resistance_variables/mt_slope_",output_resolution,".tif"), overwrite = T)
# timestamp()

slope <- raster(paste0("../R_output_files/circuitscape/resistance_variables/mt_slope_",output_resolution,".tif"))
```


## Canopy cover
```{r canopy_cover}
# 
# ## This reads in with min/max values from 0 to 255, not 0 - 100 like the NLCD legend indicates
# canopy <- raster("../R_input_files/nlcd_treecanopy/nlcd_2016_treecanopy_2019_08_31.img")
# 
# # Crop canopy layer to Montana outline extent
# canopy_crop <- crop(canopy,extent(mt_nlcd))
# 
# # Resample to be same resolution as NLCD
# canopy_rsmpl <- resample(canopy_crop,mt_nlcd)
# 
# # Mask to Montana outline
# canopy_mask <- round(mask(canopy_rsmpl,mt_nlcd),0)
# 
# # Rescale values to be from 0 to 100
# canopy_mask = canopy_mask/maxValue(canopy_mask)*100
# 
# # Plot canopy mask
# raster::plot(canopy_mask)
# 
# freq(canopy_mask)
# 
# writeRaster(canopy_mask,paste0("../R_output_files/circuitscape/resistance_variables/canopy_cover_",output_resolution,".tif"),overwrite=TRUE)

canopyCover <- raster(paste0("../R_output_files/circuitscape/resistance_variables/canopy_cover_",output_resolution,".tif"))
```

## Imperviousness
```{r imperviousness}
# # Read in NLCD Imperviousness layer
# impervious <- raster("../R_input_files/nlcd_impervious/NLCD_2016_Impervious_L48_20190405.img")
# 
# # Crop to extent of Montana shapefile
# impervious_crop <- crop(impervious,extent(mt_nlcd))
# 
# # Resample to output resolution
# impervious_rsmpl <- resample(impervious_crop,mt_nlcd)
# 
# # Mask to Montana outline
# impervious_mask <- mask(impervious_rsmpl,mt_nlcd)
# 
# # Write output file
# writeRaster(impervious_mask, paste0("../R_output_files/circuitscape/resistance_variables/imperviousness_",output_resolution,".tif"),
#             overwrite=TRUE)

imperviousness <- raster(paste0("../R_output_files/circuitscape/resistance_variables/imperviousness_",output_resolution,".tif"))
```

## Roads

### Roads Filter
```{r get_data}
# 
# # Read in Montana Transportation (Roads) shapefile
# roads = readOGR("../R_input_files/mt_transpo/Roads.shp") %>%
#   st_as_sf() %>%
#   st_simplify() %>%
#   st_transform(.,crs(nlcd))
# 
# 
# roads_arterial_avg_width <- as.data.frame(roads) %>%
#   filter(str_detect(ROADTYPE,"ARTERIAL")&ROADWIDTH > 0) %>%
#   summarise(mean_width = mean(ROADWIDTH))
# 
# ## Create a spatial data.frame of principal arterial roads
# roads_arterial_grp <- roads %>%
#   filter(str_detect(ROADTYPE,"ARTERIAL")) %>%
#   group_by(ROADTYPE) %>%
#   summarise() %>%
#   ungroup()
# 
# ## Make unique objects of roads that are not physically connected
# roads_arterial_disag = roads_arterial %>%
#   as(.,"Spatial") %>%
#   disaggregate(.) %>%
#   st_as_sf()
# 
# # ## Find area of roads
# roads_arterial_disag$rd_A <- as.numeric(st_area(roads_arterial_disag))
#  
# ## Find length of road segments. 
# roads_arterial_disag$rd_L <- as.numeric(roads_arterial_disag$rd_A/roads_arterial_avg_width$mean_width[1])/1000
# 
# # # Road Length Check
# # writeOGR(as(roads_arterial_disag,"Spatial"),
# #          dsn = "../R_output_files/scratch/",
# #          layer = "road_length_check",
# #          driver = "ESRI Shapefile",
# #          overwrite_layer = T)
# 
# ## Keep roads that are at least 1 KM long
# roads_arterial_disag_fltr <- roads_arterial_disag %>%
#   filter(rd_L >= 100) %>%
#   dplyr::select(-rd_L,-rd_A) %>%
#   group_by(ROADTYPE) %>%
#   summarise()
# 
# roads_arterial_disag_fltr = st_buffer(roads_arterial_disag_fltr,200)
# 
# ## Turn road type into a factor
# roads_arterial_disag_fltr$ROADTYPE <- as.factor(as.character(roads_arterial_disag_fltr$ROADTYPE))
# 
# ##
# roads_arterial_disag_fltr <- roads_arterial_disag_fltr %>%
#   mutate(res_value = c(70,90)) %>%
#   as(.,"Spatial")
# 
# writeOGR(roads_arterial_disag_fltr,
#          dsn = "../R_output_files/transpo/",
#          layer = "roads_arterial",
#          driver = "ESRI Shapefile",
#          overwrite_layer = T)
# 
# roads_nhs <- roads %>%
#   filter(str_detect(ROADTYPE,"NHS")) %>%
#   st_buffer(.,200) %>%
#   dplyr::select(ROADTYPE) %>%
#   mutate(res_value = 100)
# 
# nhs_ras <- rasterize(x = roads_nhs,y = mt_nlcd, field = 'res_value', fun = "first")
# arterial_ras <- rasterize(x = roads_arterial_disag_fltr, y = mt_nlcd, field = 'res_value', fun='first')
# # 
# raster::plot(nhs_ras)
# raster::plot(arterial_ras,add=T)
# # 
# road_stack <- stack(nhs_ras,arterial_ras)
# road_ras <- max(road_stack, na.rm = T)
# NAvalue(road_stack) = 0
# writeRaster(road_ras,paste0("../R_output_files/circuitscape/resistance_variables/major_roads_",output_resolution,".tif"),overwrite=T)

major_roads = raster(paste0("../R_output_files/circuitscape/resistance_variables/major_roads_",output_resolution,".tif"))
```

## Rivers
```{r rivers}
# 
# rivers_sf <- readOGR("../R_input_files/mt_hydrography/NHDArea.shp") %>%
#   st_as_sf()
# 
# rivers_fltr <- rivers_sf %>%
#   mutate(avg_width = Shape_Area/Shape_Leng) %>% # Find average width of river segment
#   filter(avg_width > 50 & Shape_Leng > 500) %>% # Keep rivers with and average width greater than 50 m  and longer than 500 m (hard to cross)
#   mutate(res_value = 50) %>% # Assign a resistance value
#   as(.,"Spatial") %>% # Turn back into a shapefile
#   spTransform(.,crs(mt_nlcd)) # Put into NLCD CRS
# 
# rivers_ras <- rasterize(rivers_fltr, mt_nlcd, field = "res_value")
# 
# writeRaster(rivers_ras, paste0("../R_output_files/circuitscape/resistance_variables/rivers_",output_resolution,".tif"), overwrite = T)

rivers <- raster(paste0("../R_output_files/circuitscape/resistance_variables/rivers_",output_resolution,".tif"))
```

## Reclassify NLCD to reflect resistances for particular Land Cover types
```{r}
## Currently for Custer County
cs_rcl_mat <- c(-Inf,10.5,NA, # NA values
                 10.5,11.5,50, # Open Water -- Natural
                 11.5,12.5,20, # Ice and Snow -- Natural
                 19,22.5,80, # Low development -- Developed
                 22.5,26,100, # High Development -- Developed
                 29,31,20, # Barren -- Natural
                 39,45,1, # Forest -- Natural
                 49,55,1, # Shrubland -- Natural
                 69,71,1, # Grassland -- Natural
                 79,85,80, # Cultivated -- Developed
                 89,95,20) # Wetlands -- Natural

nlcd_resistance <- reclassify(mt_nlcd,cs_rcl_mat)
freq(nlcd_resistance) # Use this to check that all former NLCD values were captured in the resistance reclassification (i.e. no lingering 11s for water)
writeRaster(nlcd_resistance,paste0("../R_output_files/circuitscape/resistance_variables/nlcd_resistance_rcl_",output_resolution,".tif"), overwrite=T)
```


## Putting it all together
```{r final_resistance_layers}

# Land cover
nlcd_resistance <- raster(paste0("../R_output_files/circuitscape/resistance_variables/nlcd_resistance_rcl_",output_resolution,".tif"))*10

# Roads 
major_roads <- raster(paste0("../R_output_files/circuitscape/resistance_variables/major_roads_",output_resolution,".tif"))*10

# Imperviousness 
imperviousness <- raster(paste0("../R_output_files/circuitscape/resistance_variables/imperviousness_",output_resolution,".tif"))*10

# Rivers
rivers <- raster(paste0("../R_output_files/circuitscape/resistance_variables/rivers_",output_resolution,".tif"))*10

res_stack <- stack(nlcd_resistance, major_roads, imperviousness, rivers)
max_resistance <- max(res_stack,na.rm = TRUE)

writeRaster(max_resistance,"../R_output_files/circuitscape/resistance_variables/max.tif",overwrite=T)

# Canopy Cover  
canopy_res <- raster(paste0("../R_output_files/circuitscape/resistance_variables/canopy_cover_",output_resolution,".tif"))
canopy_res_p <- canopy_res/maxValue(canopy_res)

# Slope 
slope_res = raster(paste0("../R_output_files/circuitscape/resistance_variables/mt_slope_",output_resolution,".tif"))
slope_res_p = slope_res/maxValue(slope_res)

max_res_adj <- max_resistance*(1+slope_res_p-canopy_res_p)

# Get rid of lingering roads outside of the Montana outline
resistance_mask <- mask(max_res_adj,mt_nlcd)

# Make sure all NA values are -9999 for Circuitscape
NAvalue(resistance_mask) <- -9999

# Write output ASCII
writeRaster(resistance_mask,paste0("../R_output_files/circuitscape/CS_resistance_input/montana_resistance_",output_resolution,".asc"),format="ascii",overwrite=T)
```

# Crop montana resistance to counties and write output
```{r resistance_CountyCrop}
mt_resistance = raster(paste0("../R_output_files/circuitscape/CS_resistance_input/montana_resistance_",output_resolution,".asc"))

mt_counties = read_sf("../R_input_files/mt_counties/County.shp") %>% 
  st_transform(.,crs(nlcd))

county_names = mt_counties$NAME

for(county in county_names){
  
  print(county)
  
  county_i = filter(mt_counties,
                    NAME == county)
  
   resistance_crop = raster::crop(mt_resistance,county_i)
   resistance_mask = raster::mask(resistance_crop,county_i)
  
   writeRaster(resistance_mask,
               paste0("../R_output_files/circuitscape/CS_resistance_input/",county,"_resistance_",output_resolution,".asc"),
               format = "ascii",
               overwrite=T)
   
}

```


